---
title: "About PDFs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  slidy_presentation: 
    incremental: yes
    keep_md: yes
    css: "w3c-slidy.css"
  ioslides_presentation: 
    incremental: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Four flavors of PDF

* "True", digitally created PDF
* "Image-only" PDF
* Searchable, image-based PDF
* Hybrid (image & character description) PDF

## True, perfect PDFs

Given a choice, this is the PDF you want.


The text is encoded descriptively: *This font, at this size, should be rendered in this specific location on the page*, or *canvas*. The font is described as a shape at a location, and therefore can be scaled to any size output without degradation.

<script src="https://gist.github.com/euhmeuh/61fd1beebf6d4fd5692431ed8ee1b892.js"></script>

<mark>Because each character is explicitly recorded, searches are run against a perfect index.</mark>


## "Image-only" PDFs

*Image-only* PDFs are so different from *true*, *perfect* PDFs, I wish they were called something else.

<mark>An *image-only* PDF is just a sequence of raster images--grids of color and tone values.</mark>

## "Image-only" PDFs, shortcomings

* These PDFs may include ***images of text***, but the text <mark>cannot be searched</mark>. Again, the PDF is only a sequence of images--grids of color and tone values.

* These PDFs are a sequence of, typically large, page-image files, and are therefore, frequently, <mark>too large to open and navigate easily</mark>. 

## Searchable, image-based PDFs

These PDFs are also a sequence of images, but include a hidden layer of searchable text that is, almost always, imperfect, having been created by OCR software that worked to identify characters and words by analyzing the page-images.

## Searchable, image-based PDFs, shortcomings

* Like *image-only* PDFs, these files are often <mark>too large for users to open and navigate easily</mark>.

* The OCR used to identify page-image <mark>text is error-prone</mark>. Pages typeset before computers, discolored pages, pages with schmutz, marginalia, or elaborate layouts--all of these characteristics degrade OCR accuracy, often dramatically.


## Hybrid (image & character description) PDF

Image-based PDFs with OCR-generated text can be converted to a hybrid of *true*, *perfect* and *image-based*. The software will look for OCR-generated text that the software has assigned a high confidence score, indicating it has likely identified the fonts and characters correctly, and it will throw away the page-image source and replace the image data with a described font and layout.

A single hybrid page may include encoded text that replaces part of the image, and page-image fragments for zones where the OCR confidence scores are low.

## Hybrid PDF, shortcomings

* This technique reduces the PDF size (by replacing fat image data with lean character descriptions), but it is <mark>unreliable</mark> and produces <mark>ugly</mark>, odd renderings of pages.

* Because the software discards image data that it replaces with character descriptions, you <mark>cannot recover from errors</mark> without access to the originally scanned page-images.

* Many fonts are licensed, so if Adobe cannot substitute a recognized character with the correct font, Adobe will substitute another that approximates the original.

## How does one create a *true*, *perfect* PDF from a scanned text?

Re-key the text using at least three typists and compare the three results, using discrepancies to identify errors.

## How does one create an image-based PDF with a perfect index for searching?

OCR the pages, and use special software to review potential errors and correct each one manually.

## What kind of PDFs does Imaging Services make?

Assuming a volume has text, and assuming Imaging Services has OCR software capable of identifying the language, and character set, Imaging Services creates image-based PDFs with imperfect indexes for searching.

## *END*

- [https://harvard-library-imaging-services.github.io/about_pdf/](https://harvard-library-imaging-services.github.io/about_pdf/)
- [Presentaton source file](aboutPDFs.Rmd)


```{r RUN_MANUALLY, include=FALSE, eval=FALSE}

# browseURL("https://harvard-library-imaging-services.github.io/about_pdf/")

file.copy(from = "aboutPDFs.html", to = "index.html",overwrite = TRUE)

system("git status")
system("git add --all")
system("git commit -m \"updates\"")
system("git push origin main")

```
